---
title: "B-MIS Normalization QE PelletsVesicles_ReExtractionJune2017_HILICPos_VarInjVol"
author: "Laura Hmelo"
date: "Nov 27, 2017"
output: html_document
---

#Downloaded from Ingalls Lab github 7/10/2017. Builds Sample.Key internally.  Adjusts for variable injection volumes in pooled samples. Adjusts standards in cell pellets to be equivalent to those in vesicles, since standards protocol changed in between the time each set of samples was extracted.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(ggplot2)
library(tidyr)
require(graphics); require(grDevices)
library(Hmisc)
library(gtools)
library(cowplot)
require(RColorBrewer)
# library(xlsx)
library(readr)
library(plotly)
library(stringr)
library(GGally)
library(dplyr)
```


#Setup Experiment
```{r Set experiment and fraction name,  results='hide', message=FALSE, warning=FALSE}

Location <- getwd()
Fraction <- "HILICPos"
FractionShort <- "HILICPos" #HILIC or CyanoAq or CyanoDCM
ExpDIR <-
"~/Desktop/MetabolomicsProjects/Prochlorococcus/PelletsVesiclesReExtractionJune2017"
#ExpDIR <-Location

```



#Import data from xcms - set filename here
```{r}
ExpDIR <-
"~/Desktop/MetabolomicsProjects/PelletsVesiclesReExtractionJune2017"

setwd(ExpDIR)
load("Dirs2017-06-27")

Dirs[] <- lapply(Dirs, gsub, pattern = "/Prochlorococcus", replacement = "", fixed = TRUE)



Fraction <- "HILICPos"
ResultsDIR <- as.character(Dirs[Fraction, "ResultsDIR"])
setwd(ResultsDIR)


xcms.dat <-
read.csv(paste(Fraction, "xset.filtered.csv", sep = "."), row.names = 1)
filename <- paste(Fraction, "xset.filtered.csv", sep = ".")
xcms.names <- names(xcms.dat)
```

##Read in Internal Standard data
```{r}

ISDir <- "~/Desktop/MetabolomicsProjects/PelletsVesiclesReExtractionJune2017/IS"
setwd(ISDir)


IS.dat <- read_csv("Ingalls_Lab_HILICPos_VesiclesPro_QE_Transition Results.csv")
IS.dat <- IS.dat %>%
     select(`Replicate Name`, `Precursor Ion Name`, Area) %>%
     mutate(MassFeature = `Precursor Ion Name`) %>%
  select(-`Precursor Ion Name`)

glimpse(IS.dat)
```
##Read in Internal Standard data
```{r, message=FALSE, warning=FALSE, include=FALSE}

# ISDir <- "~/Desktop/MetabolomicsProjects/Prochlorococcus/PelletsVesiclesReExtractionJune2017/IS"
# setwd(ISDir)
# IS_CyanoDCM <- read.csv("Ingalls_Lab_CyanoDCM_VesiclesPro_QE_Transition Results.csv")
# 
# IS.dat <- IS_CyanoDCM
# 
# print(xcms.names)
# 
allsamples <- xcms.names[grepl("_Smp|_Blk_2|_Blk_1|_ProcessBlk|_Poo|_Std", xcms.names)]
allsamples <- gsub('X', '', allsamples)
# IS.dat <- filter(IS.dat, Replicate.Name %in% allsamples)
# 
# 
# IS.dat <- IS.dat %>%
#      select(`Replicate.Name`, `Precursor.Ion.Name`, Area) %>%
#   rename("MassFeature" = Precursor.Ion.Name)
# glimpse(IS.dat)
```

##add bioNorm data
```{r}
SampKey_build <-  data.frame(matrix(vector(), length(allsamples), 1,
                               dimnames=list(c(), c("Replicate.Name"))),
                        stringsAsFactors=F)

SampKey <- SampKey_build %>% 
  rename(`Replicate Name` = Replicate.Name) %>%
  mutate(`Replicate Name` = allsamples) %>%
 # mutate(Sample.Group = 1) %>%
 # mutate(Treatment.Group = 1) %>%
  #mutate(Bio.Normalization = 1) %>%
  #mutate(Fraction = FractionShort) %>%
  mutate(MassFeature = "Inj_vol") %>%
  mutate(Area = ifelse(grepl("Half", `Replicate Name`),0.5, 1)) 

# setwd(ExpDIR)
# SampKey <- read_csv("sample.key.untargeted_TEMPLATE.csv") %>%
#   filter(Fraction == FractionShort) %>%
#     select(Sample.Name, Injec_vol) %>%
#   filter(!is.na(Injec_vol)) %>%
#   mutate(MassFeature = "Inj_vol") %>%
#   rename(Area = Injec_vol,
#          `Replicate Name` = Sample.Name)

IS.dat <- rbind(IS.dat, SampKey)

IS.dat.plot <- IS.dat %>% 
  filter(!MassFeature == "Inj_vol")
```


##look at extraction replication
if things don't look good with some IS make sure to fix them
or at least not include them as possibilities for normalization
```{r, echo=FALSE}
ggplot(IS.dat.plot, aes(x=`Replicate Name`, y=Area)) + geom_bar(stat="identity") + facet_wrap( ~MassFeature, scales="free_y")+theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust = 0.5), legend.position = "top")  + ggtitle("IS Raw Areas")
```

##Edit data so extraction standards are appropriately adjusted in Cell Pellets
```{r correctISErrors}

 errors <- IS.dat %>%
      filter(MassFeature=="Heavy Methionine") %>%
      filter(grepl("Pro|Std", `Replicate Name`)) %>%
      mutate(Area2 = Area) %>%
      mutate(Area= as.numeric(Area)) %>% 
      mutate(Area = Area/2)



IS.dat$Area[(IS.dat$MassFeature == "Heavy Methionine") & 
                    (grepl("Pro|Std", IS.dat$`Replicate Name`))] <- errors$Area


 errors <- IS.dat %>%
      filter(MassFeature=="Heavy Histidine") %>%
      filter(grepl("Pro|Std", `Replicate Name`)) %>%
      mutate(Area2 = Area) %>%
      mutate(Area= as.numeric(Area)) %>% 
      mutate(Area = Area/5)



IS.dat$Area[(IS.dat$MassFeature == "Heavy Histidine") & 
                    (grepl("Pro|Std", IS.dat$`Replicate Name`))] <- errors$Area

IS.dat <- IS.dat %>% 
  filter(!MassFeature == "Heavy Valine",
         !MassFeature == "Heavy Proline") %>% 
  mutate(Area = as.numeric(Area))

IS.dat.plot <- IS.dat %>% 
  filter(!MassFeature == "Inj_vol")

```
##look at extraction replication- everything better?

```{r, echo=FALSE}
q <- ggplot(IS.dat.plot, aes(x=`Replicate Name`, y=Area)) + geom_bar(stat="identity") + facet_wrap( ~MassFeature, scales="free_y")+theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust = 0.5), legend.position = "top")  + ggtitle("IS Raw Areas")
```

```{r}
print(q)
```

##Get untargeted data into long format
```{r}
IS.dat <- IS.dat %>% mutate(Replicate.Name = `Replicate Name` %>%
                                str_replace("-","."))  %>%
  select(Area, Replicate.Name, MassFeature)
replicates <- paste0("X",unique(IS.dat$Replicate.Name))

col.key <-  xcms.names %in% replicates
xcms.long <- cbind(xcms.dat[,col.key],MassFeature = xcms.dat$MassFeature) %>%
     gather(Replicate.Name, Area, -MassFeature)
```

##Rename Runs
Name structure must be:
Date_type_ID_replicate
```{r echo=FALSE}

 uniquenames<- unique(xcms.long$Replicate.Name)
 uniquenames
xcms.long <- xcms.long %>%
  mutate(Replicate.Name = Replicate.Name %>%
           str_replace("_Blk_1","_Blk_1_")%>%
           str_replace("_Blk_2","_Blk_2_") %>%
           str_replace("X1700606","1700606") %>%
           str_replace("X170606","170606"))

xcms.long <- xcms.long %>%
  mutate(Replicate.Name = Replicate.Name %>%
           str_replace("1700606", "170606"))

 uniquenames<- unique(xcms.long$Replicate.Name)
 uniquenames 

 uniquenames<- unique(IS.dat$Replicate.Name)
 uniquenames 

IS.dat <- IS.dat %>%
  mutate(Replicate.Name = Replicate.Name %>%
            str_replace("_Blk_1","_Blk_1_")%>%
           str_replace("_Blk_2","_Blk_2_") %>% 
           str_replace("170601", "170606")
         ) 

IS.dat <- IS.dat %>%
  mutate(Replicate.Name = Replicate.Name %>%
            str_replace("1700606", "170606"))

uniquenames<- unique(IS.dat$Replicate.Name)
 uniquenames 

```

##Calculate mean values for each IS
```{r ISmeans}
IS.means <- IS.dat %>% filter(!grepl("Blk", Replicate.Name)) %>%
  mutate(MassFeature = as.factor(MassFeature))%>%
     group_by(MassFeature) %>%
     summarise(ave = mean(Area))
```

##Normalize to each internal Standard
Actually scaling/adjusting by dividing by IS peak area and multiplying by the average IS peak area.
Output is a dataframe 'area.norm' that has each peak adjusted
to each possible internal standards
##This step can take a pretty long time
```{r echo=FALSE}
binded <- rbind(IS.dat, xcms.long)
wArea<- binded %>% 
     select(Replicate.Name, MassFeature, Area) %>%
     spread(key=MassFeature, value=Area) %>%
     as.data.frame

IS.list <- unique(IS.dat$MassFeature)
this.IS <- IS.list[1]
area.norm <- wArea[,-1] %>% 
          sapply(FUN = function(x) x/wArea[,grep(this.IS,
                                                 names(wArea))]) %>%
          as_data_frame %>% mutate(Replicate.Name = wArea$Replicate.Name) %>%
          gather(MassFeature,Area_Norm, -Replicate.Name)
this.mean <- IS.means %>% filter(MassFeature==this.IS) %>% 
     select(ave) %>% as.numeric
area.norm <- area.norm %>% mutate(Area_Norm = Area_Norm*this.mean)
key <- ncol(area.norm)
count <- length(which(!is.na(area.norm$Area_Norm)))/
               length(unique(area.norm$Replicate.Name))
names(area.norm)[key] <- paste(this.IS,"Norm.Area",sep=".")
# print(paste(1, this.IS, count, sep="-"))

for (i in 2:length(IS.list)){
     this.IS <- IS.list[i]
     if(length(wArea[,grep(this.IS, names(wArea))])!=0){
          this.norm <- wArea[,-1] %>% 
               sapply(FUN = function(x) x/wArea[,grep(this.IS,
                                                      names(wArea))]) %>%
               as_data_frame %>% 
               mutate(Replicate.Name = wArea$Replicate.Name) %>%
               gather(MassFeature,Area_Norm, -Replicate.Name)
          this.mean <- IS.means %>% filter(MassFeature==this.IS) %>% 
               select(ave) %>% as.numeric
          this.norm <- this.norm %>% mutate(Area_Norm = Area_Norm*this.mean)
          key <- ncol(area.norm)
          area.norm[,key+1] <- this.norm$Area_Norm
          names(area.norm)[key+1] <- paste(this.IS,"Norm.Area",sep=".")
          count <- length(which(!is.na(this.norm$Area_Norm)))/
               length(unique(this.norm$Replicate.Name))
          # print(paste(i, this.IS, count, sep="-"))
     }
}
#glimpse(area.norm)
write_csv(area.norm, paste(Fraction, ".area.norm.csv", sep=""))
```


#Break Up the Names
Name structure must be:
Date_type_ID_replicate
```{r echo=FALSE}
mydata_new <- area.norm %>% separate(Replicate.Name, 
                                      c("runDate",
                                        "type","SampID","replicate"),"_") %>%
     mutate(Run.Cmpd = paste(area.norm$Replicate.Name,area.norm$MassFeature))
binded <- binded %>% mutate(Run.Cmpd = paste(Replicate.Name, MassFeature))

dat <- full_join(binded, mydata_new)
#glimpse(dat)
write.csv(dat, paste(Fraction, ".dat.temp.csv", sep=""))
```

##Compare normalizations
get mean, sd, and rsd for each normalization of each compound for each sample type.
Output is a dataframe called 'rsd.stats'
```{r}
no.blank.dat <- dat %>% filter(type =="Smp" | type =="Poo") 

rsd.stats <- no.blank.dat %>% select(-Replicate.Name,  -Run.Cmpd,
                                     -runDate, -replicate) %>%
     gather(Normer, Value, -MassFeature, -type, - SampID) %>%
     group_by(MassFeature, type, SampID, Normer) %>%
     summarise(m = mean(Value, na.rm=T), 
                sd = sd(Value, na.rm=T), rsd = sd/m)
write.csv(rsd.stats, paste(Fraction, ".rsd.stats.csv", sep=""))
#glimpse(rsd.stats)
```

##Cleanup RSD data and add a minimizing column 
One based on the multiple injetions of the pooled sample "PooModel"
and one based on the injections of the biological replicates of samples "SmpModel"

# [1] missing in github version of SmpModel code. Won't run without it.
```{r newModeMaker}

setwd(ExpDIR)
rsd.stats <-read.csv(paste(Fraction, ".rsd.stats.csv", sep=""), row.names = 1)
rsd.clean <- rsd.stats %>% filter(!is.na(m))  %>%
     filter(Normer!="Area")

# SmpModel <- rsd.clean %>% filter(type=="Smp") %>%
#      select(-m, -sd) %>%
#      group_by(MassFeature, Normer) %>%
#      summarise(Mean.rsd = mean(rsd, na.rm=T)) %>%
#      summarise(SmpModelRSD = min(Mean.rsd))


SmpModel <- rsd.clean %>% filter(type=="Smp") %>%
     select(-m, -sd) %>%
     group_by(MassFeature, Normer) %>%
     dplyr::summarise(Mean.rsd = mean(rsd, na.rm=T)) %>%
     dplyr::summarise(SmpModelRSD = min(Mean.rsd),
               Smp.Picked.IS = unique(Normer)[which.min(Mean.rsd)][1])

PooModel <- rsd.clean %>% filter(type=="Poo") %>%
     select(-m, -sd) %>%
     group_by(MassFeature, Normer) %>%
     dplyr::summarise(Mean.rsd = mean(rsd, na.rm=T)) %>%
     dplyr::summarise(PooModelRSD = min(Mean.rsd),
               Poo.Picked.IS = unique(Normer)[which.min(Mean.rsd)][1])
Models <- full_join(PooModel, SmpModel)
#glimpse(Models)
```

##Merge Models and RSD data all together
```{r}
rsd.total <- full_join(rsd.stats, Models)
```

##Adjust the pooled model --> PooPlus model aka B-MIS model
Uses the PooModel for all compounds detected in the Pooled samples.
If a compound was detected in the sample but not the pooled then use the SmpModel.
Names this new complied model "PooPlus" (aka B-MIS)
Output of this section is 'rsd.total': A dataframe with mean, sd, and rsd for each compound, sample type, and normalization combination and a column indicating the poo, smp and pooPlus (B-MIS) internal standard selections 
```{r echo=FALSE}
rsd.total <- rsd.total %>%
     mutate(PooPlus.IS = Poo.Picked.IS) %>%
     mutate(PooPlusModelRSD = PooModelRSD)

no.poo.rows <-which(is.na(rsd.total$PooModelRSD))
rsd.total$PooPlus.IS[no.poo.rows] <- rsd.total$Smp.Picked.IS[no.poo.rows]
rsd.total$PooPlusModelRSD[no.poo.rows] <- rsd.total$SmpModelRSD[no.poo.rows]

rsd.total <- rsd.total %>% 
     select(-PooModelRSD,-PooPlusModelRSD,-SmpModelRSD)
#glimpse(rsd.total)
```

##Write out RSD data, if you want
```{r, eval=FALSE}
#write.csv(rsd.total, "All_normalization_rsd_dat.csv")
```


##Get Actual Normalized data not just the rsds
Only keep PooPlus (BMIS model)
```{r, echo=FALSE}
setwd(ExpDIR)
dat <- read.csv(paste(Fraction, "dat.temp.csv", sep="."), row.names = 1)
# 
# models <- rsd.total %>% ungroup %>%
#      select(MassFeature, PooPlus.IS) %>%
#      group_by(MassFeature) %>%
#      summarise(PooPlusModel.IS = unique(PooPlus.IS))
# dat <- dat %>% filter(!is.na(Replicate.Name))
# dat.join <- as.data.frame(full_join(dat, models)) %>%
#   mutate(PooPlusModel = NA)
# 
# split.on.IS <- as.factor(dat.join$PooPlusModel.IS)
# split.dat.join <- split(dat.join, split.on.IS)
# for (i in 1:length(split.dat.join)){
#   col.key <-  which(names(split.dat.join[[i]])==names(split.dat.join)[i])
#   split.dat.join[[i]]$PooPlusModel <- split.dat.join[[i]][,col.key]
# }
# 
# unsplit.dat.join <- unsplit(split.dat.join, split.on.IS)
# 
# ##write out that full normalized data
# ##The output of the B-MIS normalized data is in the column named PooPlusModel.
# new.filename <- paste("BMISd",filename,sep="_")
# # write.csv(unsplit.dat.join, paste(as.character(Dirs[Fraction, "ResultsDIR"]), "/", new.filename, sep = "", collapse = NULL))

```


#Rebuilt this last chunk from older TQS BMIS script.  Takes hours to run for untargeted data (let run overnight)
```{r}
models <- rsd.total %>% ungroup %>%
     select(MassFeature,
            Poo.Picked.IS,Smp.Picked.IS,PooPlus.IS) %>%
     group_by(MassFeature) %>%
     dplyr::summarise(PooModel.IS = unique(Poo.Picked.IS),
               SmpModel.IS = unique(Smp.Picked.IS),
               PooPlusModel.IS = unique(PooPlus.IS))
dat <- dat %>% filter(!is.na(Replicate.Name))
dat.join <- full_join(dat, models)

dat.join <- dat.join %>%
  mutate(PooModel.IS = gsub(' ', '.', PooModel.IS)) %>%
  mutate(SmpModel.IS = gsub(' ', '.', SmpModel.IS)) %>%
  mutate(PooPlusModel.IS = gsub(' ', '.', PooPlusModel.IS)) 

starttime <- Sys.time()

PooModel <- c()
SmpModel <-c()
PooPlusModel <- c()
for (i in 1:nrow(dat.join)){
     if(!is.na(dat.join$PooModel.IS[i])){
          PooModel <- c(PooModel,dat.join[i,dat.join$PooModel.IS[i]])
     } else {
          PooModel <- c(PooModel,NA)
     }
     if(!is.na(dat.join$SmpModel.IS[i])){
          SmpModel <- c(SmpModel,dat.join[i,
                                            dat.join$SmpModel.IS[i]])
     } else {
          SmpModel <- c(SmpModel,NA)
     }
     if(!is.na(dat.join$PooPlusModel.IS[i])){
          PooPlusModel <- c(PooPlusModel,dat.join[i,
                                            dat.join$PooPlusModel.IS[i]])
     } else {
          PooPlusModel <- c(PooPlusModel,NA)
     }
  #print(i)
}
dat.join$PooModel <- PooModel
dat.join$SmpModel <- SmpModel
dat.join$PooPlusModel <- PooPlusModel
dat.join <- dat.join %>% unnest(PooModel) %>% unnest(SmpModel) %>%
     unnest(PooPlusModel)
glimpse(dat.join)

endtime <- Sys.time()
totaltime <- endtime - starttime
```

##Write out that full normalized data
The output of the B-MIS normalized data is in the column named PooPlusModel.
```{r, eval=FALSE}
new.filename <- paste("BMISed_data",filename,sep="_")
write.csv(dat.join, new.filename)
```
