---
title: "DataExploration_CyanoDCM_June2017Data"
author: "Laura Hmelo"
date: "December 05, 2017"
output: html_document
---

```{r setup, include=FALSE}

library(stringr)
library(ggplot2)
library(tidyverse)
library(reshape2)
library(cowplot)

```


```{r setup, include=FALSE}

Experiment <- "CyanoDCM_June2017Data"

#define key directories
ExpDIR <-
"~/Desktop/MetabolomicsProjects/PelletsVesiclesReExtractionJune2017"
functionDIR <- "~/Desktop/MetabolomicsProjects/RScripts"
targetcompoundpath <- "~/Desktop/MetabolomicsProjects/RScripts/data"

#load in Params and relevant Dirs for Experiment
setwd(ExpDIR)
Params <- read.csv("Params.csv",  row.names = 1)
load("Dirs2017-06-27")

Dirs[] <- lapply(Dirs, gsub, pattern = "/Prochlorococcus", replacement = "", fixed = TRUE)

#Define fraction list (may be reassigned within chunks)
FractionList <- c("HILICNeg", "HILICPos", "CyanoAq", "CyanoDCM")


#Define experimental groups
Treat1ID <- "Pro9313"
Treat2ID <- "Vesicle9313"





```


#load in files for data exploration using just BMIS-normalized-CyanoDCM fraction 

**CHANGE INPUT FILE TO BMIS NORMALIZED FILE!**
```{r setup CyanoDCM}


Fraction <- "CyanoDCM"
ResultsDIR <- as.character(Dirs[Fraction, "ResultsDIR"])
setwd(ResultsDIR)
#NeutralMassList <- read.csv(paste(Fraction, "NeutralMassList.csv", sep="."))
xset.filtered <- read.csv(paste(Fraction, ".xset.filtered.csv", sep=""), row.names=1)
xset.normal <- read.csv(paste(Fraction, "PelletsVesicles.9313.xset.vesnormalized.csv", sep="."))
xset.biovolumenormal <- read.csv(paste(Fraction, "PelletsVesicles.9313.xset.biovolumenormalized.csv", sep="."))


#load("sig.groupnames.RData")

names <- colnames(xset.filtered)
Samples <- names[grepl("_Smp_", names)]
Blanks <- names[grepl("Blk", names)]
Pooled <- names[grepl("Poo", names)]

Treat1 <- Samples[grepl(as.character(Treat1ID), Samples)]
Treat2 <- Samples[grepl(as.character(Treat2ID), Samples)]
```

```{r}
xset.mzRT <- xset.filtered %>% select(MassFeature, mz, RT)
xset.normal <- xset.normal %>% left_join(xset.mzRT, "MassFeature")
xset.biovolumenormal <- xset.biovolumenormal %>% left_join(xset.mzRT, "MassFeature")


```


#Generate MFs list for mass features most prominent in 9313 or 9212 within CyanoDCM fraction
```{r CyanoDCM eicplotter, include=FALSE}
#MFs for all mass features where 9313 is at least 4x 9212

# setwd(functionDIR)
# source("EICPlots_2Groups.function.R")
# 
# eicplotter(MFs_9312high_short, "CyanoDCM", "CyanoDCM_ProPellets_MFs_9312_75MostDifferent")
# eicplotter(MFs_9313high_short, "CyanoDCM", "CyanoDCM_ProPellets_MFs_9313_75MostDifferent")
```

#Plotting EICs for internal standards to pdf for three fractions
```{r isplotter all fractions, echo=FALSE}
# setwd(functionDIR)
# source("ISplotter.function.R")
# FractionList <- c("HILICPos", "CyanoAq", "CyanoDCM")
# isplotter()
```

#Plotting EICs for targetted metabolites to pdf for three fractions
```{r tcplotter all fractions, message=FALSE, warning=FALSE, include=FALSE}
# setwd(functionDIR)
# source("TCplotter.function.R")
# FractionList <- c("HILICPos", "CyanoAq", "CyanoDCM")
# tcplotter()

```

#raw data plotted for all fractions
```{r}
# 
# setwd(functionDIR)
# source("logplot.function.R")
# FractionList <- c( "HILICNeg", "HILICPos", "CyanoAq", "CyanoDCM")
# logplot()
# 
#Mass features on positive Y axis are enriched in Treat1: Cell Pellets
```
#Plot raw CyanoDCM data
```{r plot CyanoDCM vesnorm}

Fraction <- "CyanoDCM"
ResultsDIR <- as.character(Dirs[Fraction, "ResultsDIR"])
setwd(ResultsDIR)
data <-
xset.filtered
 
  
    
    # names <- colnames(data)
    # Samples <- names[grepl("Smp", names)]
    # 
    # Treat1 <- Samples[grepl(as.character(Treat1ID), Samples)]
    # Treat2 <- Samples[grepl(as.character(Treat2ID), Samples)]

 data_test <- data %>%
    filter((AveVesicles9313 > (AveBlk1 * 2) & 
             AveVesicles9313 > (AveBlk2 * 2) & 
             (AveVesicles9312 + Vesicles9312.sd) > (AveBlk1 - Blk1.sd) &
             (AveVesicles9312 + Vesicles9312.sd) > (AveBlk2 - Blk2.sd) &
             (AvePro9312 + Pro9312.sd) > (AveBlk2 - Blk2.sd) &
             (AvePro9312 + Pro9312.sd) > (AveProcessBlk - ProcessBlk.sd) &
             (AvePro9313 + Pro9313.sd) > (AveBlk2 - Blk2.sd) &
             (AvePro9313 + Pro9313.sd) > (AveProcessBlk - ProcessBlk.sd)) |
             
             (AveVesicles9312 > (AveBlk1 * 2) & 
                AveVesicles9312 > (AveBlk2 * 2) & 
                (AveVesicles9313 + Vesicles9313.sd) > (AveBlk1 - Blk1.sd) &
                (AveVesicles9313 + Vesicles9313.sd) > (AveBlk2 - Blk2.sd) &
                (AvePro9312 + Pro9312.sd) > (AveBlk2 - Blk2.sd) &
                (AvePro9312 + Pro9312.sd) > (AveProcessBlk - ProcessBlk.sd) &
                (AvePro9313 + Pro9313.sd) > (AveBlk2 - Blk2.sd) &
                (AvePro9313 + Pro9313.sd) > (AveProcessBlk - ProcessBlk.sd)) |
  
             
             (AvePro9313 > (AveProcessBlk * 2) & 
                AvePro9313 > (AveBlk2 * 2) & 
                (AveVesicles9312 + Vesicles9312.sd) > (AveBlk1 - Blk1.sd) &
                (AveVesicles9312 + Vesicles9312.sd) > (AveBlk2 - Blk2.sd) &
                (AveVesicles9313 + Vesicles9313.sd) > (AveBlk1 - Blk1.sd) &
                (AveVesicles9313 + Vesicles9313.sd) > (AveBlk2 - Blk2.sd) &
                (AvePro9312 + Pro9312.sd) > (AveBlk2 - Blk2.sd) &
                (AvePro9312 + Pro9312.sd) > (AveProcessBlk - ProcessBlk.sd)) |
             
             (AvePro9312 > (AveProcessBlk * 2) & 
                AvePro9312 > (AveBlk2 * 2) & 
                (AveVesicles9312 + Vesicles9312.sd) > (AveBlk1 - Blk1.sd) &
                (AveVesicles9312 + Vesicles9312.sd) > (AveBlk2 - Blk2.sd) &
                (AveVesicles9313 + Vesicles9313.sd) > (AveBlk1 - Blk1.sd) &
                (AveVesicles9313 + Vesicles9313.sd) > (AveBlk2 - Blk2.sd) &
                (AvePro9313 + Pro9313.sd) > (AveBlk2 - Blk2.sd) &
                (AvePro9313 + Pro9313.sd) > (AveProcessBlk - ProcessBlk.sd)) 
           )
             

```

#Plot CyanoDCM data normalized to vesicle and cell counts and internal standards
```{r plot CyanoDCM vesnorm}

Fraction <- "CyanoDCM"
ResultsDIR <- as.character(Dirs[Fraction, "ResultsDIR"])
setwd(ResultsDIR)


data <- xset.normal %>% 
  filter(AveTreat1 > 1 & AveTreat2 > 1)
    
    
    
    
    Treat1ID <- "Pro9313"
    Treat2ID <- "Vesicle9313"
    
    names <- colnames(data)
    Samples <- names[grepl("Smp", names)]
    
    Treat1 <- Samples[grepl(as.character(Treat1ID), Samples)]
    Treat2 <- Samples[grepl(as.character(Treat2ID), Samples)]
    
    data$AveTreat1<- rowMeans(data[, Treat1])
    data$AveTreat2<- rowMeans(data[, Treat2])
  
    
    data <- data %>%
      mutate(Treat1.sd = apply(data[,Treat1], 1, sd)) %>%
      mutate(Treat2.sd = apply(data[,Treat2], 1, sd)) %>%
      mutate(Treat1.percent.sd = ((Treat1.sd/AveTreat1)*100)) %>%
      mutate(Treat2.percent.sd = ((Treat2.sd/AveTreat2)*100))
      
    
    data$log2ratio <-
      ifelse(
        data$AveTreat1 > data$AveTreat2,
        log2(data$AveTreat1 / data$AveTreat2),
        ifelse(
          data$AveTreat1 < data$AveTreat2,
          -(log2(
            data$AveTreat2 / data$AveTreat1
          )),
          "other"
        )
      )
    
    #replacing Inf with "5" in order to treat the column as numeric
    data$log2ratio <- gsub('Inf', '25', data$log2ratio)
    data$log2ratio <- as.numeric(data$log2ratio)
    
    Treatsdf <- data[, c(Treat1,Treat2) ]
     data$pvalue <- apply(Treatsdf, 1, function(x) {t.test(x[1:3],x[4:6])$p.value})
     
     data$pvalue.wt <- apply(Treatsdf, 1, function(x) {wilcox.test(x[1:3],x[4:6], correct=FALSE)$p.value})
     
     
     
    sig.data <- subset(data, data$pvalue < 0.05)
    
    
    
   p <- ggplot(sig.data, aes(x = mz, y = log2ratio, color = pvalue)) +
      geom_point() +
      scale_color_continuous(low = "red", high = "blue") +
      labs(title = (paste(Fraction, Experiment, "Normalized to cell and vesicle count", sep = " "))) +
      theme_classic()
    
 print(p)
 

setwd(ResultsDIR)
 write.csv(data, paste(Fraction, "logplotdata", "bionorm", "ISreq", "csv", sep="."))
save_plot("CyanoDCM_PelletsVsVesicles9313_NormalizedToCellsVesicles.pdf", p,  base_aspect_ratio = 2)
```

```{r}
  p <- ggplot(data %>% filter(log2ratio < -1), aes(x = mz, y = RT, size = log2ratio, color = log2ratio)) +
      geom_point() +
      scale_color_continuous(low = "red", high = "blue") +
      labs(title = (paste(Fraction, Experiment, "Normalized to cell and vesicle count", sep = " "))) +
    scale_size_continuous(range=c(4,2))+
      theme_classic()
    
 print(p)
 

# setwd(ResultsDIR)
#  write.csv(sig.data, paste(Fraction, "logplotdata", "bionorm", "csv", sep="."))
# save_plot("CyanoDCM_PelletsVsVesicles9313_NormalizedToCellsVesicles.pdf", p,  base_aspect_ratio = 2)
```


#Plot CyanoDCM vesicles and cells normalized to biovolume, IS required
```{r plot CyanoDCM normalized to biovolume, IS required}
data <-
      xset.biovolumenormal
    
    Treat1ID <- "Pro9313"
    Treat2ID <- "Vesicle9313"
    
    names <- colnames(data)
    Samples <- names[grepl("Smp", names)]
    
    Treat1 <- Samples[grepl(as.character(Treat1ID), Samples)]
    Treat2 <- Samples[grepl(as.character(Treat2ID), Samples)]
    
    data$AveTreat1<- rowMeans(data[, Treat1])
    data$AveTreat2<- rowMeans(data[, Treat2])
    Treat1.sd <- apply(data[,Treat1], 2, sd)
    # data_test <- data %>% mutate(Treat1.sd = apply(data[,Treat1], 2, sd)) %>% mutate(Treat2.sd = apply(data[,Treat2], 2, sd))
    
    data$log2ratio <-
      ifelse(
        data$AveTreat1 > data$AveTreat2,
        log2(data$AveTreat1 / data$AveTreat2),
        ifelse(
          data$AveTreat1 < data$AveTreat2,
          -(log2(
            data$AveTreat2 / data$AveTreat1
          )),
          "other"
        )
      )
    
    #replacing Inf with "5" in order to treat the column as numeric
    data$log2ratio <- gsub('Inf', '25', data$log2ratio)
    data$log2ratio <- as.numeric(data$log2ratio)
    
    Treatsdf <- data[, c(Treat1,Treat2) ]
     data$pvalue <- apply(Treatsdf, 1, function(x) {t.test(x[1:3],x[4:6])$p.value})
     
        data$pvalue.wt <- apply(Treatsdf, 1, function(x) {wilcox.test(x[1:3],x[4:6], correct=FALSE)$p.value})
     
     
     
     
    sig.data <- subset(data, data$pvalue < 0.05)
    
    
    
   p <- ggplot(sig.data, aes(x = mz, y = log2ratio, color = pvalue)) +
      geom_point() +
      scale_color_continuous(low = "red", high = "blue") +
      labs(title = (paste(Fraction, Experiment, "Normalized to biovolume", sep = " "))) +
      theme_classic()
    
 print(p)
 

setwd(ResultsDIR)
 write.csv(data, paste(Fraction, "logplotdata", "biovolume","ISreq", "csv", sep="."))
save_plot("CyanoDCM_PelletsVsVesicles9313_NormalizedToBiovolume.pdf", p,  base_aspect_ratio = 2)

```
