---
title: "PelletsVesiclesALLStrains_June2017"
author: "Laura Hmelo"
date: "June 27, 2017"
output:
---

#Vesicles were extracted in January 2017 and rerun with Cell Pellets, Freshly extracted in May 2017.

#Set filepaths
```{r Set filepaths}

ExpDIR <- "~/Desktop/MetabolomicsProjects/PelletsVesiclesReExtractionJune2017" #raw data path
functionDIR <- "~/Desktop/MetabolomicsProjects/RScripts" #location of scripts
#outputpath <- "~/Desktop/MetabolomicsProjects/XCMS/EddyQE141216" #path that files are written to 
targetcompoundpath <- "~/Desktop/MetabolomicsProjects/RScripts/data" #this is where MRM_Methods_Table.csv is stored
```

#Set parameters
```{r Set parameters and define groups}
Experiment <- "VesiclesVsPellets9313"
setwd(ExpDIR)
Params <- read.csv("Params.csv",  row.names=1)

FractionList <- c("CyanoAq", "CyanoDCM", "HILICNeg", "HILICPos")

pvalue <- 0.05

#Pick 2 IDs that will work to call Biological replicates to compared for the QC step
Treat1ID <- "Pro9313"
Treat2ID <- "Pro9312"
Treat3ID <- "Vesicle9313"
Treat4ID <- "Vesicle9312"

```


```{r Set your Directories, include=FALSE}
#ExpDIR <-paste(basepath, Experiment, sep = "/", collapse = NULL)
FractionDIR <- c()
ResultsDIR <- c()
DataDIR <- c()
for (j in 1:length(FractionList)){
  FractionDIR[j] <- paste(ExpDIR, "/", FractionList[j], sep = "", collapse = NULL)
  ResultsDIR[j] <- paste(FractionDIR[j], "/Results", Sys.Date(), sep = "", collapse = NULL)
 dir.create(file.path(ResultsDIR[j])) #, showWarnings = FALSE) #Comment out if you don't need to make the folder i.e. you've already run XCMS and need to redo the next steps
  DataDIR <-  paste(FractionDIR, "/rawdat", sep = "", collapse = NULL)
  }
Dirs <- data.frame(FractionList, FractionDIR, ResultsDIR, DataDIR)
#Dirs <- Dirs %>% mutate(ResultsDIR = ResultsDIR %>%  str_replace("2017-01-24","2017-01-22") ) #If you are accessing results made on previous days, use this bit of code (^) to redirect to older results
rownames(Dirs) <- Dirs$FractionList
setwd(ExpDIR)
save(Dirs, file=paste("Dirs", Sys.Date(), sep=""))
```

```{r recreating Dirs}
# FractionDIR <- c()
# ResultsDIR <- c()
# DataDIR <- c()
# for (j in 1:length(FractionList)){
#   FractionDIR[j] <- paste(ExpDIR, "/", FractionList[j], sep = "", collapse = NULL)
#   ResultsDIR[j] <- paste(FractionDIR[j], "/Results", Sys.Date(), sep = "", collapse = NULL)
#  #dir.create(file.path(ResultsDIR[j])) #, showWarnings = FALSE) #Comment out if you don't need to make the folder i.e. you've already run XCMS and need to redo the next steps
#   DataDIR <-  paste(FractionDIR, "/rawdat", sep = "", collapse = NULL)
#   }
# Dirs <- data.frame(FractionList, FractionDIR, ResultsDIR, DataDIR)
# Dirs <- Dirs %>% mutate(ResultsDIR = ResultsDIR %>%  str_replace("2017-02-24","2017-02-22") ) #If you are accessing results made on previous days, use this bit of code (^) to redirect to older results
# rownames(Dirs) <- Dirs$FractionList
# setwd(ExpDIR)
# save(paste(Dirs, "2017-02-22", sep=""))
```

#Run XCMSfunction

```{r Run XCMSfunction on all fractions specified in FractionList, include=FALSE}
setwd(functionDIR)
source("XCMS.V2.function.R")

xcmsfunction()


```
#*******THIS FUNCTION MUST BE CUSTOMIZED FOR EACH DATA SET!!!!!!!******


```{r, include=FALSE}
setwd(ExpDIR)
load("Dirs2017-06-27")
Dirs[] <- lapply(Dirs, gsub, pattern = "/Prochlorococcus", replacement = "", fixed = TRUE)
Params <- read.csv("Params.csv",  row.names = 1)

setwd(functionDIR)
source("mfmaker_VesiclesAndPellets_4Treats.function.R")
mfmaker_ttest()

#done 6-28-17
#Rerun 12-12-2017 after updating mfmaker function
```

#Extract internal standards using skyline, completed 7-05-17

#RedoB-MIS, completed 12-12-2017
#Used Script: B-MIS-Normalization_QE_PelletsVesiclesReExtractionJune2017_CyanoDCM_VarInjVol.Rmd
#Output: BMISed_data_CyanoDCM.xset.filtered.csv

#Run Camera_Function
```{r Run CAMERA, include=FALSE}
#Calculate adducts and isotopes for mass features and calculate neutral mass for ions with two or more adducts
#
library(tidyverse)
library(stringr)

setwd(ExpDIR)
load("Dirs2017-06-27")

Params <- read.csv("Params.csv",  row.names = 1)

Dirs <- Dirs %>% 
  mutate(ResultsDIR = ResultsDIR %>%  str_replace("/Prochlorococcus","") ) %>%
   mutate(DataDIR = DataDIR %>%  str_replace("/Prochlorococcus","") ) %>%
   mutate(ResultsDIR = ResultsDIR %>%  str_replace("/Prochlorococcus","") ) %>%
   mutate(FractionDIR = FractionDIR %>%  str_replace("/Prochlorococcus","") )
  
row.names(Dirs) <- Dirs[, "FractionList"]

j=1
Fraction <- FractionList[j]
    print(Fraction)
    ResultsDIR <- as.character(Dirs[Fraction, "ResultsDIR"])
    setwd(ResultsDIR)
   load(paste(Fraction, "xset3.RData", sep="."))
   
   save(xset3, file=paste(Fraction, "xset3", "original", "RData", sep = "."))
   
   old_filepaths <- filepaths(xset3) 
   
   new_filepaths <- gsub('/Prochlorococcus', '', old_filepaths)
   
   filepaths(xset3) <- new_filepaths
    setwd(ResultsDIR)
    save(xset3, file=paste(Fraction, "xset3", "RData", sep = "."))
    
  
   
setwd(functionDIR)
source("CAMERA.V2.function.R")
camera (PPM = 5)

```




```{r, include=FALSE}
setwd(functionDIR)
source("qcplotter.V2_0628.function.R")
qcplotter()

```

#Dereplication of XCMS output
```{r Dereplicate XCMS output, include=FALSE}
#these can be targeted compounds or standards.  Right now code is set up to accomodate RP or HILIC targetted compounds
#There don't appear to be any matches within Vesicles, CyanoAq or CyanoDCM.

setwd(functionDIR)
source("TargetCompounds.V2.function.R")
#setwd(outputpath)
mfmatch(X,Y=xset.allpeaks, column, 5,0.4)

```


#Run Camera post-processing
#Load in CAMERA xsAnnotate object and Peaklist

```{r}
setwd(ExpDIR)
load("Dirs2017-02-22")
Params <- read.csv("Params.csv",  row.names = 1)

```


```{r Run all CAMERA post-processing, include=FALSE}
setwd(functionDIR)
source("CAMERApostprocessing.V2.function.R")
camerapostprocess()


```


```{r}
#Neutral Mass List function
#Compare mass feature priority lists generated by XCMS script (e.g. 50 most abundant compounds, compounds with p<0.05) to adduct list.  Dereplicate priority lists and output a list of neutral masses relevant to data set.  Option to output a list of "pseudo-neutral masses" calculated by subtracting an H+ from all mass features which have no associated adduct.

#Requires OtherIons, *.filtered peaklist from XCMS script, and sig.groupnames from XCMS script
setwd(functionDIR)
source("NeutralMassList.V2.function.R")
NMLfunction() 
```
```{r, message=FALSE, warning=FALSE, include=FALSE}
setwd(ExpDIR)
Params <- read.csv("Params.csv",  row.names=1)
Dirs <- read.csv("Dirs2017-02-03.csv", row.names=1)
```

#Plot TICs with generic rainbow colors
```{r}
# setwd(functionDIR)
# source("getTICs.V2.function.R")
# 
# getTICs(rt="corrected")

```


#Plot TICs with lines colored by sample type
```{r, include=FALSE}
setwd(ExpDIR)
load("Dirs2017-06-27")
  
  
setwd(functionDIR)
source("getTICs.4Treats.function.R")

getTICs(rt="corrected")
#done 6-28-17
```

